Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes.

### 2020-04-13 Ben_Manuscripts/hdphmm/Draft.tex
add.
Added missing intro figs

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath} % or simply amstext
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\newcommand{\angstrom}{\textup{\AA}}
\newcommand{\colormap}{jet}  % colorbar to use
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}
\usepackage{xr}
\usepackage{setspace}
\doublespacing

\externaldocument[S-]{Supporting_Information}

@@

  % BJC2: I'm not putting a lot of effort into the introduction since I'll only write
  % an abbreviated form for my dissertation. I'll work on the intro after my defense.  
  %MRS7: sounds good, I won't look at this today.

@@

  In our previous 
%MRS:work
study,
  we designed two different approaches which used
  solute time series in order to parameterize stochastic models that could be used
  to project transport on much longer timescales. In our first approach we modeled 
  solute trajectories as subordinated fractional Brownian and L\'evy motion, called
  the anomalous diffusion (AD) model. We generated solute trajectories by generating
  a series of anti-correlated hops separated by random periods of entrapment drawn 
  from a power law distribution. Our second approach treated solute motion as a Markov
  state model with state-dependent dynamics, called the Markov state-dependent dynamical
  model (MSDDM). We parameterized the state transition probabilities between
  each of eight discrete states as well as the solute dynamics within each of these
  states. We generated stochastic trajectory realizations by drawing a state
  sequence based on the transition probability matrix and incorporating the state dynamics
  while solutes were trapped in each state.

@@

  In this work we attempt to minimize human effort by applying the infinite hidden 
  Markov Model (IHMM), a nonparameteric Bayesian classification technique, in order to 
  automatically detect and infer the parameters of an unknown number of latent
  autoregressive (AR) modes present in solute center-of-mass time series trajectories.
  We aim to address two primary scientific questions with this work. First, we want
  to know if the IHMM algorithm can help researchers efficiently uncover underlying
  transport mechanisms which give rise to different solute dynamical behavior.
  Second, we want to evaluate the ability of our model to reproduce solute trajectories
  in a stochastic manner so that we can use it to predict macroscopic transport 
  properties which would otherwise require much longer MD simulations.

@@

  Because independent solute trajectories generated by MD simulations share 
  similar state dynamics, we can use the IHMM to generate stochastic trajectory 
  realizations which sample from all behavior exhibited by MD. We verify that our
  stochastic realizations reproduce the expected hopping and trapping behavior 
  exhibited by MD. We quantitatively validate these trajectories by comparing 
  their mean squared displacement (MSD) to MSDs measured by MD.

@@

  We ran all MD simulations and energy minimizations using GROMACS 2018. Our 
  python implementation of the IHMM algorithm is available online at 
  \texttt{https://github.com/bencoscia/hdphmm}. All other post-simulation 
  trajectory analysis tools are available online at \texttt{https://github.com/shirtsgroup/LLC\_Membranes}.

@@

  We let each solute undergo 5 $\mu$s of MD simulation. We used a time step of 2 fs
  at a pressure of 1 bar and temperature of 300K controlled by the Parrinello-Rahman 
  barostat~\cite{???} and velocity rescale thermostat~\cite{???} respectively. We recorded configurations every 0.5 ns.

@@

  Hidden Markov models (HMMs) are a useful and widely used technique for modeling
  sequences of observations where the probability of the next observation in a 
  sequence depends, at least in part, on a previous unobserved, latent or hidden,
  state.~\cite{beal_infinite_2002} In the context of our simulations, the observations
  correspond to the center of mass coordinates of the solutes versus time, and the
  states correspond to the dynamical behavior which give rise to those types
  of observations. The probability of transitioning to a state based on the current
  state is mathematically defined in terms of an $n\times n$ transition probability
  matrix, $T$, where $n$ is the number of states. Unfortunately, standard HMMs 
  require $n$ to be known \textit{a priori}. One can partially overcome this by 
  testing a range of numbers of hidden states and determining which is the best 
  representation of the data.

  %BJC: I'm not sure how much of this I should include. This is pretty much 
  % a reproduction of Fox et al's explanation: https://ieeexplore.ieee.org/abstract/document/5563110
  The infinite-state HMM overcomes this drawback by placing a hierarchical
  Dirichlet process (HDP) prior on the transition probabilities. Using some 
  base probability distribution, $H$, a Dirichlet process (DP) generates discrete 
  distributions, $G_0$, over a countably infinite number of probability measures:
  \begin{equation}
      G_0 = \sum_{k=1}^{\infty} \beta_k \delta_{\theta_k} ~~ \theta_k \sim H, \beta \sim GEM(\gamma)
  \end{equation}
  where the $\theta_k$ are values drawn from the base distribution and the weights
  $\beta_k$ come from a stick-breaking process parameterized by the concentration 
  parameter $\gamma$ (equivalently referred to as GEM($\gamma$)). In shorthand, this
  specification can be expressed as $G_0 \sim DP(\gamma, H)$. The concentration 
  parameter, $\gamma$, expresses one's confidence in the base distribution $H$.
  We use a uniform base distribution for $H$. When $\gamma\to 0$, the first 
  weight of $G_0$, $\beta_1$, approaches unity and for $\gamma\to\infty$, the weights
  become uniform and $G_0$ closely resembles $H$. Each row, $G_j$, of the transition 
  matrix is produced by drawing from a DP specified using the $\beta$ vector as a 
  discrete base distribution and a separate concentration parameter, $\alpha$.
  \begin{equation}
      G_j = \sum_{k=1}^{\infty} \pi_{jk} \delta_{\theta_k} ~~ \pi_j \sim DP(\alpha, \beta)
  \end{equation}
  This hierarchical specification ensures that the transition probabilities in 
  each row share the same support points \{$\theta_1$, ..., $\theta_k$\}.
  Once the model has converged only a finite number of states will have significant
  sampling.

@@

  % BJC2: It might be useful to walk through the following procedure with an example 
  % trajectory
  % TODO in the SI

  %Based on the algorithms developed by Fox et al.~\cite{fox_sticky_2007}, 
  Using the IHMM framework, we estimated the most likely number and sequence of hidden
  states in each solute center-of-mass trajectory while simultaneously inferring VAR(1)
  parameters for each state and the overall state transition probability matrix, $T$.
  We created a python implementation of this process which we heavily adapted from
  the MATLAB code of Fox et al.~\cite{fox_sticky_2007} Parameter estimation is iterative. 
  Therefore, we looked for convergence of each entry in the $\mathbf{c}$ vector as well
  as the $A$ and $\Sigma$ matrices of Equation~\ref{eqn:var}. We detected equilibration 
  of the parameters using the module \texttt{pymbar.timeseries.detect\_equilibration} on
  the time series of parameter estimates. Plots illustrating parameter convergence are 
  available in Section~\ref{S-section:convergence} of the Supporting Information. We refer the interested reader to 
  much more extensive descriptions of the inference and sampling procedures used to 
  estimate the VAR(1) parameters and the state sequence. ~\cite{beal_infinite_2002,teh_hierarchical_2006,van_gael_beam_2008,fox_nonparametric_2009,fox_bayesian_2010}

@@

  We start by applying the IHMM algorithm to the 3D solute center-of-mass coordinate 
  trajectories transformed relative to the closest pore center. We tracked the solute's
  motion along the pore axis with the center-of-mass $z$ coordinate. Using the nearest
  pore center as the origin, we represented the radial distance of each solute's 
  center-of-mass from the pore center in 2 dimensions, $x$ and $y$ 
  %(see Figure~\ref{fig:cartesian_cylinder}). 
  By working in Cartesian coordinates, we avoid
  mathematical complexity introduced by cylindrical coordinates while estimating the 
  state sequence.

@@

  We applied the IHMM to each of the 24 solute trajectories independently.
  Although the IHMM is capable of identifying an infinite number of states, 
  a Dirichlet process tends to exhibit a ``rich get richer" effect, favoring
  a fewer number of states. By applying the algorithm to each trajectory 
  independently, we reduce the possibility of lumping together multiple 
  similar states which we would prefer to stay separated before clustering.
  Note that the states identified by the IHMM are heavily influenced by 
  the Gaussian prior placed on $\mathbf{c}$ in Equation~\ref{eqn:var}. 
  In Section~\ref{S-section:prior_guesses} of the Supporting Information we outline our method for 
  choosing reliable prior parameters. The entries of $A$ and $\Sigma$ do 
  not vary over a wide range, so the final parameters were relatively insensitive
  to their priors.

  We ran 2000 iterations of the IHMM procedure in order to arrive at converged 
  state sequences and parameters for each state. Most parameters converged 
  quickly (within 50-100 iterations) while others took up to 1000 iterations.
  %BJC6: still need to make these figures.
  We defined the finalized parameters of each state as the average of the 
  parameters from each iteration recorded after the equilibration time point.
  We found the equilibration time point of each component of the $\mathbf{c}$ 
  vector as well as $A$ and $\Sigma$ matrices, and then used the longest 
  equilibration time of all dimensions as the equilibration time point.
  Further details are available in Section~\ref{S-section:convergence} of
  the Supporting Information.

@@

  We reduced the parameter space used agglomerative clustering, a hierarchical
  clustering approach which uses a linkage criteria in order to successively merge
  similar clusters until a desired intracluster distance threshold or number of
  clusters is reached. We used the Ward linkage criteria, which works to minimize 
  the sum of the squared differences within all clusters.
    %MRS6: explain why Ward (make the cluster as minimally disperse?)
   	%BJC6: I'll add some plots to the SI and then write a better justification for
	% the paper. This is lower priority in mind for the thesis.
  We elected to choose the number of clusters rather than the distance threshold.
  For our data, non-parametric methods such as Bayesian Gaussian mixture models tend
  to delocalize the clusters in parameter space (see Section~\ref{S-section:agglomerative}
  of the Supporting Information).

@@

  Choosing the number of clusters can be somewhat subjective so
  we attempted to add some structure to the selection process by following
  a set of qualitative and quantitative guidelines. We used the silhouette 
  test in order to score the quality of clustering as a function of the number 
  of clusters chosen. For our data, the silhouette test generally favors the 
  lowest number of clusters possible (see Section~\ref{S-section:nclusters} of
  the Supporting Information). However, choosing too few clusters tends to 
  not distinguish between visually obvious differences in 
  dynamic behavior. This results in finalized parameter sets that are averages of
  distinct behavior which presents further problems with the predictive modeling
  that we discuss later on. We aimed to maintain the highest silhouette score,
  and thus lowest number of clusters, possible while verifying that visually 
  distinct states stayed separated. Based on these guidelines, we decided to 
  group the $A$ and $\Sigma$ parameters into five clusters each and the radial
  means into three clusters. Of the 75 distinct states allowed by this formulation,
  the solutes in this study showed behavior from 32--36 distinct clusters. While
  this may seem like a high number of states, many are sampled infrequently.

@@

  We used the parameter inference component of the IHMM algorithm in order to 
  infer $A$ and $\Sigma$ of the clustered states.
   %BJC2: I'm unsure of the precise reasoning here.
  We could not simply take the mean of the clustered $A$ and $\mathbf{e}_t$ 
  parameters because it is not clear that this is a linear 
  operation for this problem. To circumvent this problem, we modified the ($r$,
  $z$) solute trajectories so that they had a mean of zero, leaving only the 
  fluctuations. We did this by the mean of each same-state segment of the 
  unclustered trajectory. We used the IHMM algorithm on this modified trajectory
  to infer the clustered state parameters by fixing the clustered state sequence.

@@

  We estimated the lifetime of hydrogen bonds by recording the length of 
  sequential frames where solutes remained hydrogen bonded. We still counted sequences
  where hydrogen bonds were broken for a single frame before reforming. Consistent
  with our previous work, we reported the 95th percentile of hydrogen bond lifetimes
  since their distribution is not Gaussian and to emphasize longer trapping periods.

@@

  We developed a way to measure the local density of the membrane at arbitrary
  points in the unit cell. We histogrammed the positions in three dimensions
  Since our system is in a monoclinic unit cell, we periodically replicated
  the system in the +/- x, y and z directions and then chose the bounds on the
  histogram in order to create a rectangular box encompassing the unit cell with
  a 1 nm buffer between the histogram and unit cell boundaries. We then used a regular
  grid interpolator in order to allow interpolation at arbitrary points within the grid.

  We quantified solute motion using the time averaged mean squared displacement (MSD)
  of the solute center of mass trajectories. The time-averaged MSD measures all observed
  displacements over time lag $\tau$:
  \begin{equation}
  	\overline{z^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (z(t + \tau) - z(t))^2 dt
  \label{eqn:tamsd}
  \end{equation}
  where T is the length of the trajectory. We reported 1$\sigma$ confidence intervals 
  based on the results of 200 bootstrap trials. For each trail, we calculated the mean of 
  $n$ random trajectories chosen with replacement from the pool of trajectories, where 
  $n$ is the number of trajectories.

@@

  %MRS7: more descriptive title?  What about methanol?
  \subsubsection*{The Case of Methanol}

  %MRS6: likely just the summary of the methanol behavior should go in the main paper - the rest should be SI.
  Methanol exhibits a range of dynamic behavior. In Figure~\ref{fig:mechanism_map}, we
  plot a representative 2D methanol trajectory in cylindrical coordinates. The coordinates
  are colored according to the cluster with which methanol's dynamics are most consistent,
  which implies that this particular methanol trajectory exhibits 8 distinct dynamical 
  behaviors (in order of their first appearance, they are colored dark yellow, red, light
  blue, light green, orange, blue, teal and lime green). The solute time series are coupled
  with color coded bars which describe the solute's local environment and physical 
  interactions that it undergoes at that trajectory frame. From the top, they describe 
  whether the solute is associated with sodium, whether the solute is hydrogen bonding
  with the monomer as well as to what part of the monomer it is hydrogen bonding, and 
  finally, the local number density surrounding the solute.

@@

  It is clear that there is a wealth of information in these plots, but they don't give
  the complete picture needed in order to formulate LLC membrane design principles. One
  can analyze each segment sequentially and form a clear picture of this instance of 
  methanol's behavior without the need to visualize the trajectory with software like VMD.
  However, it is still possible to gain a more general understanding of methanol's average
  behavior.

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.6\textwidth}
  \includegraphics[width=\textwidth]{common_states_MET.pdf}
  \caption{}\label{fig:common_states_MET_lines}
  \end{subfigure}
  \begin{subfigure}{0.35\textwidth}
  % BJC: I'll bootstrap some error bars on this by drawing the transition probabilities 
  % from a Dirichlet distribution.
  \includegraphics[width=\textwidth]{A_sigma_scatter_MET.pdf}
  \caption{}\label{fig:A_sigma_scatter_MET}
  \includegraphics[width=\textwidth]{dwell_times_MET.pdf}  %BJC5: maybe better as 95 % confidence rather than mean
  \caption{}\label{fig:dwell_times_MET}
  \end{subfigure}
  \caption{We can learn the most about methanol's behavior by studying the dynamics
  of the states common to many of the methanol trajectories. In (a), we show representative
  dynamics of six states that are common to more than one third of the solute trajectories. 
  The $y$-axis labels of the $r$ coordinate plot specify the radial mean of each clustered 
  state. 
  %MRS7: it's confusing to have the R not in order of r. Emphasize this more in the caption, maybe think of something a bit different for the final paper. Reads odd to have numbers not in order along an axis.
The percentage of trajectories in which the states appear are labeled to the right of
  the $r$ plots. States 1--5 appear in the trajectory shown in Figure~\ref{fig:mechanism_map}
  and are colored to match for easy visual comparison. All time series have a mean of zero and are
  shifted to put the mean in line with the $y$-axis labels for the purpose of visualizing them. 
  (b) We can begin to understand the solute behaviors demonstrated in (a) by understanding their VAR 
  parameters. States 1 and 2 have higher covariances in both the axial and radial directions compared
  to all other states. This should lead to significant solute displacement. Covariance in each dimension
  appears to have a strong correlation. States 3--6 all have relatively low covariances, which may
  be characteristic of trapped states. They are primarily distinguished based on the magnitude of their 
  autoregressive coefficients, A, and radial means. Higher values of A imply motion which wanders 
  more about the mean. States 1, 3 and 5 have radial means near 1 nm or less, which suggest that they
  occur close to the pore centers. States 2, 4 and 6 occur well within the tail region of the membrane.
  (c) Using the transition probabilities (p), we estimated the average time spent within each 
  of the states using 1000 simulations of a Bernoulli process where we continued to draw until transitioning
  out of the state with probability 1 - p. Of the prevalent states, states 1 and 
  2 exhibit the shortest dwell times alongside their high covariances. States 5 and 7 have the longest dwell times in the pore and tail regions respectively.
  }\label{fig:common_states_MET}
  \end{figure}

@@

  We applied the IHMM to the other three solutes in this study: urea, acetic acid
  and ethylene glycol. 
  \begin{itemize}  
    \item Not only can we use the model to study the dynamics of each solute 
    individually, but we can compare their behavior in order to improve our
    understanding of more subtle mechanistic details.
  \end{itemize}

  The size and chemical functionality of the solutes dictates the interactions which
  influence solute transport. 
  \begin{itemize}  
    \item In Figure~\ref{fig:hbonds_assoc_summary}, we show that the solutes 
    donate hydrogen bonds and associate with sodium ions to varying degrees.
    \item We also show the solute MSDs after a 1000 ns time lag. 
    \item It is not immediately obvious why the MSDs follow this trend.
    \item The states identified by the IHMM can help us shed some light.
  \end{itemize}

>>>>

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath} % or simply amstext
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\newcommand{\angstrom}{\textup{\AA}}
\newcommand{\colormap}{jet}  % colorbar to use
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}
\usepackage{xr}
\usepackage{setspace}
\doublespacing
%\DeclareUnicodeCharacter{03C9}{$\omega$}
\externaldocument[S-]{Supporting_Information}

@@

@@

  In our previous study, we designed two different approaches which used
  solute time series in order to parameterize stochastic models that could be used
  to project transport on much longer timescales. In our first approach we modeled 
  solute trajectories as subordinated fractional Brownian and L\'evy motion, called
  the anomalous diffusion (AD) model. We generated solute trajectories by generating
  a series of anti-correlated hops separated by random periods of entrapment drawn 
  from a power law distribution. Our second approach treated solute motion as a Markov
  state model with state-dependent dynamics, called the Markov state-dependent dynamical
  model (MSDDM). We parameterized the state transition probabilities between
  each of eight discrete states as well as the solute dynamics within each of these
  states. We generated stochastic trajectory realizations by drawing a state
  sequence based on the transition probability matrix and incorporating the state dynamics
  while solutes were trapped in each state.

@@

  In this work we attempt to minimize human effort by applying the infinite hidden 
  Markov Model (IHMM)~\cite{beal_infinite_2002,calderon_inferring_2015}, a nonparameteric 
  Bayesian classification technique, in order to automatically detect and infer the 
  parameters of an unknown number of latent autoregressive (AR) modes present in solute
  center-of-mass time series trajectories.
  We aim to address two primary scientific questions with this work. First, we want
  to know if the IHMM algorithm can help researchers efficiently uncover underlying
  transport mechanisms which give rise to different solute dynamical behavior.
  Second, we want to evaluate the ability of our model to reproduce solute trajectories
  in a stochastic manner so that we can use it to predict macroscopic transport 
  properties which would otherwise require much longer MD simulations.

@@

  Because independent solute trajectories generated by MD simulations share 
  similar state dynamics, we can use the IHMM to generate stochastic trajectory 
  realizations which sample from all behavior exhibited by MD. We verify that our
  stochastic realizations reproduce the expected hopping and trapping behavior 
  exhibited by MD. We quantitatively validate these trajectories by comparing 
  their mean squared displacement (MSD) to those measured by MD.

@@

  We ran all MD simulations and energy minimizations using GROMACS 2018.
  ~\cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}  
  Our python implementation of the IHMM algorithm is available online at 
  \texttt{https://github.com/bencoscia/hdphmm}. All other post-simulation 
  trajectory analysis tools are available online at \texttt{https://github.com/shirtsgroup/LLC\_Membranes}.

@@

  We let each solute undergo 5 $\mu$s of MD simulation. We used a time step of 2 fs
  at a pressure of 1 bar and temperature of 300K controlled by the Parrinello-Rahman 
  barostat~\cite{parrinello_polymorphic_1981} and velocity rescale 
  thermostat~\cite{bussi_canonical_2007} respectively. We recorded configurations every 0.5 ns.

@@

  Hidden Markov models (HMMs) are a useful and widely used technique for modeling
  sequences of observations where the probability of the next observation in a 
  sequence depends, at least in part, on a previous unobserved, latent or hidden,
  state.~\cite{beal_infinite_2002} In the context of our simulations, the observations
  correspond to the center of mass coordinates of the solutes versus time, and the
  states correspond to the dynamical behavior which give rise to those types
  of observations. The probability of transitioning to a state based on the current
  state is mathematically defined in terms of an $n\times n$ transition probability
  matrix, $T$, where $n$ is the number of states. Unfortunately, standard HMMs 
  require $n$ to be known \textit{a priori}. One can partially overcome this by 
  testing a range of numbers of hidden states and determining which is the best 
  representation of the data.~\cite{pohle_selecting_2017}

  %BJC: I'm not sure how much of this I should include. This is pretty much 
  % a reproduction of Fox et al's explanation: https://ieeexplore.ieee.org/abstract/document/5563110
  The infinite-state HMM overcomes this drawback by placing a hierarchical
  Dirichlet process (HDP) prior on the transition probabilities.~\cite{fox_bayesian_2010} Using some 
  base probability distribution, $H$, a Dirichlet process (DP) generates discrete 
  distributions, $G_0$, over a countably infinite number of probability measures:
  \begin{equation}
      G_0 = \sum_{k=1}^{\infty} \beta_k \delta_{\theta_k} ~~ \theta_k \sim H, \beta \sim GEM(\gamma)
  \end{equation}
  where the $\theta_k$ are values drawn from the base distribution and the weights
  $\beta_k$ come from a stick-breaking process parameterized by the concentration 
  parameter $\gamma$ (equivalently referred to as GEM($\gamma$)).~\cite{halmos_random_1944} In shorthand, this
  specification can be expressed as $G_0 \sim DP(\gamma, H)$. The concentration 
  parameter, $\gamma$, expresses one's confidence in the base distribution $H$.
  We use a uniform base distribution for $H$. When $\gamma\to 0$, the first 
  weight of $G_0$, $\beta_1$, approaches unity and for $\gamma\to\infty$, the weights
  become uniform and $G_0$ closely resembles $H$. Each row, $G_j$, of the transition 
  matrix is produced by drawing from a DP specified using the $\beta$ vector as a 
  discrete base distribution and a separate concentration parameter, $\alpha$.
  \begin{equation}
      G_j = \sum_{k=1}^{\infty} \pi_{jk} \delta_{\theta_k} ~~ \pi_j \sim DP(\alpha, \beta)
  \end{equation}
  This hierarchical specification ensures that the transition probabilities in 
  each row share the same support points \{$\theta_1$, ..., $\theta_k$\}.
  Once the model has converged only a finite number of states will have significant
  sampling.

@@

  %Based on the algorithms developed by Fox et al.~\cite{fox_sticky_2007}, 
  Using the IHMM framework, we estimated the most likely number and sequence of hidden
  states in each solute center-of-mass trajectory while simultaneously inferring VAR(1)
  parameters for each state and the overall state transition probability matrix, $T$.
  We created a python implementation of this process which we heavily adapted from
  the MATLAB code of Fox et al.~\cite{fox_sticky_2007} Parameter estimation is iterative. 
  Therefore, we looked for convergence of each entry in the $\mathbf{c}$ vector as well
  as the $A$ and $\Sigma$ matrices of Equation~\ref{eqn:var}. We detected equilibration 
  of the parameters using the module \texttt{pymbar.timeseries.detect\_equilibration}~\cite{chodera_simple_2016} on
  the time series of parameter estimates. Plots illustrating parameter convergence are 
  available in Section~\ref{S-section:convergence} of the Supporting Information. We refer the interested reader to 
  much more extensive descriptions of the inference and sampling procedures used to 
  estimate the VAR(1) parameters and the state sequence. ~\cite{beal_infinite_2002,teh_hierarchical_2006,van_gael_beam_2008,fox_nonparametric_2009,fox_bayesian_2010}

@@

  We start by applying the IHMM algorithm to the 3D solute center-of-mass coordinate 
  trajectories transformed relative to the closest pore center. We tracked the solute's
  motion along the pore axis with the center-of-mass $z$ coordinate. Using the nearest
  pore center as the origin, we represented the radial distance of each solute's 
  center-of-mass from the pore center in 2 dimensions, $x$ and $y$.
  %(see Figure~\ref{fig:cartesian_cylinder}). 
  By working in Cartesian coordinates, we avoid
  mathematical complexity introduced by cylindrical coordinates while estimating the 
  state sequence.

@@

  We applied the IHMM to each of the 24 solute trajectories independently.
  Although the IHMM is capable of identifying an infinite number of states, 
  a Dirichlet process tends to exhibit a ``rich get richer" effect, favoring
  a fewer number of states.~\cite{dreyer_discovering_2011} By applying the algorithm to each trajectory 
  independently, we reduce the possibility of lumping together multiple 
  similar states which we would prefer to stay separated before clustering.
  Note that the states identified by the IHMM are heavily influenced by 
  the Gaussian prior placed on $\mathbf{c}$ in Equation~\ref{eqn:var}. 
  In Section~\ref{S-section:prior_guesses} of the Supporting Information we outline our method for 
  choosing reliable prior parameters. The entries of $A$ and $\Sigma$ do 
  not vary over a wide range, so the final parameters were relatively insensitive
  to their priors.

  We ran 2000 iterations of the IHMM procedure in order to arrive at converged 
  state sequences. We evaluated the convergence of the state sequence based on
  the convergence of the VAR parameters. The parameters of each states
  generally converge within 500--1000 iterations. However, the boundaries of 
  each state segment fluctuate which can lead to high variance in the $A$
  and $\Sigma$ parameters in segments with a relatively low number of emissions.
  In these cases, the mean in each dimension tends to be a more reliable indicator of
  convergence (see Figure~\ref{S-fig:convergence3d} of the Supporting Information).

  We defined the finalized parameters of each state as the average of the 
  parameters from each iteration recorded after the equilibration time point.
  We found the equilibration time point of each component of the $\mathbf{c}$ 
  vector as well as $A$ and $\Sigma$ matrices, and then used the longest 
  equilibration time of all dimensions as the equilibration time point.
%  Further details are available in Section~\ref{S-section:convergence} of
%  the Supporting Information.

@@

  We reduced the parameter space used agglomerative clustering, a hierarchical
  clustering approach which uses a linkage criteria in order to successively merge
  similar clusters until a desired intracluster distance threshold or number of
  clusters is reached.~\cite{pedregosa_scikit-learn_2011} We used the Ward linkage criteria, which works to minimize 
  the sum of the squared differences within all clusters.~\cite{ward_hierarchical_1963}
    %MRS6: explain why Ward (make the cluster as minimally disperse?)
   	%BJC6: I'll add some plots to the SI and then write a better justification for
	% the paper. This is lower priority in mind for the thesis.
  We elected to choose the number of clusters rather than the distance threshold.
  For our data, non-parametric methods such as Bayesian Gaussian mixture models~\cite{pedregosa_scikit-learn_2011,gelman_bayesian_2013} tend
  to delocalize the clusters in parameter space (see Section~\ref{S-section:agglomerative}
  of the Supporting Information).

@@

  Choosing the number of clusters can be somewhat subjective so
  we attempted to add some structure to the selection process by following
  a set of qualitative and quantitative guidelines. We used the silhouette 
  test in order to score the quality of clustering as a function of the number 
  of clusters chosen.~\cite{kaufman_finding_2009} For our data, the silhouette test generally favors the 
  lowest number of clusters possible (see Section~\ref{S-section:nclusters} of
  the Supporting Information). However, choosing too few clusters tends to 
  not distinguish between visually obvious differences in 
  dynamic behavior. This results in finalized parameter sets that are averages of
  distinct behavior which presents further problems with the predictive modeling
  that we discuss later on. We aimed to maintain the highest silhouette score,
  and thus lowest number of clusters, possible while verifying that visually 
  distinct states stayed separated. Based on these guidelines, we decided to 
  group the $A$ and $\Sigma$ parameters into five clusters each and the radial
  means into three clusters. Of the 75 distinct states allowed by this formulation,
  the solutes in this study showed behavior from 32--36 distinct clusters. While
  this may seem like a high number of states, many are sampled infrequently.

@@

  We used the parameter inference component of the IHMM algorithm in order to 
  infer $A$ and $\Sigma$ of the clustered states.
   %BJC2: I'm unsure of the precise reasoning here.
  We could not simply take the mean of the clustered $A$ and $\mathbf{e}_t$ 
  parameters because it is not clear that this is a linear 
  operation for this problem. To circumvent this problem, we modified the ($r$,
  $z$) solute trajectories so that they had a mean of zero, leaving only the 
  fluctuations. We did this by subtracting the mean from each same-state segment of the 
  unclustered trajectory. We used the IHMM algorithm on this modified trajectory
  to infer the clustered state parameters by fixing the clustered state sequence.

@@

  We estimated the lifetime of hydrogen bonds by recording the length of 
  sequential frames where solutes remained hydrogen bonded. We still counted sequences
  where hydrogen bonds were broken for a single frame before reforming. Consistent
  with our previous work, we reported the 95th percentile of hydrogen bond lifetimes
  since their distribution is not Gaussian and to emphasize longer trapping periods.~\cite{coscia_chemically_2019}

@@

  We developed a way to measure the local density of the membrane at arbitrary
  points in the unit cell. We histogrammed the positions in three dimensions
  Since our system is in a monoclinic unit cell, we periodically replicated
  the system in the +/- x, y and z directions and then chose the bounds on the
  histogram in order to create a rectangular box encompassing the unit cell with
  a 1 nm buffer between the histogram and unit cell boundaries.~\cite{van_der_walt_numpy_2011} We then used a regular
  grid interpolator in order to allow interpolation at arbitrary points within the grid.~\cite{virtanen_scipy_2020}

  We quantified solute motion using the time averaged mean squared displacement (MSD)
  of the solute center of mass trajectories.~\cite{meroz_toolbox_2015} The time-averaged MSD measures all observed
  displacements over time lag $\tau$:
  \begin{equation}
  	\overline{z^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (z(t + \tau) - z(t))^2 dt
  \label{eqn:tamsd}
  \end{equation}
  where T is the length of the trajectory. We reported 1$\sigma$ confidence intervals 
  based on the results of 200 bootstrap trials. For each trail, we calculated the mean of 
  $n$ random trajectories chosen with replacement from the pool of trajectories, where 
  $n$ is the number of trajectories.

@@

  \subsubsection*{The Dynamics of Methanol}

  Methanol exhibits a range of dynamic behavior. In Figure~\ref{fig:mechanism_map}, we
  plot a representative 2D methanol trajectory in cylindrical coordinates. The coordinates
  are colored according to the cluster with which methanol's dynamics are most consistent,
  which implies that this particular methanol trajectory exhibits eight distinct dynamical 
  behaviors (in order of their first appearance, they are colored dark yellow, red, light
  blue, light green, orange, blue, teal and lime green). The solute time series are coupled
  with color coded bars which describe the solute's local environment and physical 
  interactions that it undergoes at that trajectory frame. From the top, they describe 
  whether the solute is associated with sodium, whether the solute is hydrogen bonding
  with the monomer as well as to what part of the monomer it is hydrogen bonding, and 
  finally, the local number density surrounding the solute.

@@

  It is clear that there is a wealth of information in these plots, but they don't give
  the complete picture needed in order to formulate LLC membrane design principles. One
  can analyze each segment sequentially and form a clear picture of this instance of 
  methanol's behavior without the need to visualize the trajectory with software like VMD.~\cite{humphrey_vmd:_1996}
  However, it is still possible to gain a more general understanding of methanol's average
  behavior.

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.6\textwidth}
  \includegraphics[width=\textwidth]{common_states_MET.pdf}
  \caption{}\label{fig:common_states_MET_lines}
  \end{subfigure}
  \begin{subfigure}{0.35\textwidth}
  % BJC: I'll bootstrap some error bars on this by drawing the transition probabilities 
  % from a Dirichlet distribution.
  \includegraphics[width=\textwidth]{A_sigma_scatter_MET.pdf}
  \caption{}\label{fig:A_sigma_scatter_MET}
  \includegraphics[width=\textwidth]{dwell_times_MET.pdf}  %BJC5: maybe better as 95 % confidence rather than mean
  \caption{}\label{fig:dwell_times_MET}
  \end{subfigure}
  \caption{We can learn the most about methanol's behavior by studying the dynamics
  of the states common to many of the methanol trajectories. In (a), we show representative
  dynamics of six states that are common to more than one third of the solute trajectories. 
  The $y$-axis labels of the $r$ coordinate plot specify the radial mean of each clustered 
  state. 
  %MRS7: it's confusing to have the R not in order of r. Emphasize this more in the caption, 
  %maybe think of something a bit different for the final paper. Reads odd to have numbers not in order along an axis.
  %BJC7: indeed
  The percentage of trajectories in which the states appear are labeled to the right of
  the $r$ plots. States 1--5 appear in the trajectory shown in Figure~\ref{fig:mechanism_map}
  and are colored to match for easy visual comparison. All time series have a mean of zero and are
  shifted to put the mean in line with the $y$-axis labels for the purpose of visualizing them. 
  (b) We can begin to understand the solute behaviors demonstrated in (a) by understanding their VAR 
  parameters. States 1 and 2 have higher covariances in both the axial and radial directions compared
  to all other states. This should lead to significant solute displacement. Covariance in each dimension
  appears to have a strong correlation. States 3--6 all have relatively low covariances, which may
  be characteristic of trapped states. They are primarily distinguished based on the magnitude of their 
  autoregressive coefficients, A, and radial means. Higher values of A imply motion which wanders 
  more about the mean. States 1, 3 and 5 have radial means near 1 nm or less, which suggest that they
  occur close to the pore centers. States 2, 4 and 6 occur well within the tail region of the membrane.
  (c) Using the transition probabilities (p), we estimated the average time spent within each 
  of the states using 1000 simulations of a Bernoulli process where we continued to draw until transitioning
  out of the state with probability 1 - p. Of the prevalent states, states 1 and 
  2 exhibit the shortest dwell times alongside their high covariances. States 5 and 7 have the longest dwell times in the pore and tail regions respectively.
  }\label{fig:common_states_MET}
  \end{figure}

@@

  We applied the IHMM to the other three solutes in this study: urea, acetic acid
  and ethylene glycol. Not only can we use the model to study the dynamics of each 
  solute individually, but we can compare their behavior in order to improve our
  understanding of more subtle mechanistic details.

  The size and chemical functionality of the solutes dictates the interactions which
  influence solute transport. In Figure~\ref{fig:hbonds_assoc_summary}, we show that
  the solutes donate hydrogen bonds and associate with sodium ions to varying degrees.
  We also show the solute MSDs after a 1000 ns time lag. It is not immediately obvious
  why the MSDs follow this trend. The states identified by the IHMM shed
  some light on the reasoning.
